---
title: "experience similarity and accurate interpersonal understanding"
author: "Yoona Kang"
date: "`r Sys.Date()`"
output: html_document
 html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen=999)
```


# load packages
```{r}
library("lme4")
library("lmerTest")
library("ggplot2")
library("Rcpp")
library("lm.beta")
library("psych")
library(jtools)
```

#load and inspect data
```{r}
df <- read.csv('~/Box Sync/Varela_Data_822363/data/cleaned/varela_10_21_2020.csv')
df$cond <- relevel(df$cond, ref='control')
colnames(df)[colnames(df)=="Q56"] <- "edu"

table(scale(df$EA_corr)) #2 outliers
table(scale(df$factual_accuracy)) #no outlier
table(scale(df$speaker_similar_exp))#no outlier
table(scale(df$maas))#no outlier
df2 <- subset(df, EA_corr > .5) #excluding two outliers

```


#demographics
```{r}
summary(df$age)
sd(df$age)
table (df$race)
```

#MAAS crombach's alpha
```{r}
df_maas <- df[ , c("maas1", "maas2", "maas3", "maas4", "maas5", "maas6", "maas7",
                      "maas8", "maas9","maas10", "maas11","maas12","maas13", "maas14","maas15")]    
alpha(df_maas)
```


# define function to make results table
```{r}
make_table = function(model) {
  require(broom)
  require(knitr)
  
  print(data.frame(R2 = summary(model)$r.squared,
             adjusted_R2 = summary(model)$adj.r.squared,
             df = summary(model)$df[2]) %>%
    knitr::kable(format = "pandoc", digits = 3))
    
  print(model %>%
    broom::tidy(conf.int = TRUE) %>%
    mutate(p.value = ifelse(p.value < .001, "< .001", gsub("0.(.*)", ".\\1", sprintf("%.3f", p.value))),
           `95% CI` = sprintf("%.3f, %.3f", conf.low, conf.high),
           term = gsub("as.factor", "", term)) %>%
    select(-contains("conf")) %>%
    rename("b" = estimate,
           "SE" = std.error,
           "t" = statistic,
           "p" = p.value) %>%
    select(term, b, `95% CI`, everything()) %>%
    knitr::kable(format = "pandoc", digits = 3))
}
```

#Results

#past experience --> factual accuracy
```{r}
test <- lm(factual_accuracy ~speaker_similar_exp+cond,df)
make_table(test)
```
#past experience --> empathic accuracy
```{r}
test <-lm(EA_corr ~speaker_similar_exp+cond,df)
make_table(test)
```

#excluding outliers
```{r}
test <-lm(EA_corr ~speaker_similar_exp+cond,df2)
make_table(test)
```
#mindfulness * experience similarity --> factual accuracy
```{r}
test <-lm(factual_accuracy ~maas*speaker_similar_exp+cond,df)
make_table(test)
```

#mindfulness * experience similarity --> factual accuracy
```{r}
test <-lm(EA_corr ~ maas*speaker_similar_exp +cond,df)
make_table(test)
#excluding outliers
test <-lm(EA_corr ~ maas*speaker_similar_exp +cond,df2)
make_table(test)
```

#simple slopes analysis
```{r}
print(mass_low<- round((mean(df$maas) - sd(df$maas)),1))
print(mass_mid<- round((mean(df$maas)),1))
print(mass_high <- round((mean(df$maas) + sd(df$maas)),1))

#center variables
df$maas_cen = df$maas - mean(df$maas,na.rm=T)
df$exp_cen = df$speaker_similar_exp- mean(df$speaker_similar_exp,na.rm=T)
df$EA_corr_cen = df$EA_corr - mean(df$EA_corr,na.rm=T)

#+-sd for mindfulness
df$maas_low <- df$maas_cen + sd(df$maas_cen, na.rm=T)
df$maas_high <- df$maas_cen - sd(df$maas_cen, na.rm=T)

test <- lm(EA_corr_cen~ exp_cen*maas_high+as.factor(cond), df)
make_table(test)
test <- lm(EA_corr_cen ~ exp_cen*maas_cen +as.factor(cond), df)
make_table(test)
test <- lm(EA_corr_cen ~ exp_cen*maas_low +as.factor(cond), df)
make_table(test)
```


#mindfulness --> empathic accuracy at different levels of experience similarity
```{r}
df$exp_low <- df$exp_cen + sd(df$exp_cen, na.rm=T)
df$exp_high <- df$exp_cen - sd(df$exp_cen, na.rm=T)

test <- lm(EA_corr_cen~ maas*exp_low+as.factor(cond), df)
make_table(test)
test <- lm(EA_corr_cen~ maas*exp_cen+as.factor(cond), df)
make_table(test)
test <- lm(EA_corr_cen~ maas*exp_high+as.factor(cond), df)
make_table(test)

```


#Figures
```{r}
##Figure 1

df$Mindfulness <- df$maas
test <- lm(EA_corr~ Mindfulness * speaker_similar_exp +as.factor(cond), df)
interact_plot(test, pred = speaker_similar_exp, modx = Mindfulness,
              x.label = "Experience similarity", y.label = "Empathic accuracy", interval=T, int_type="confidence") 

interact_plot(test, pred = Mindfulness, modx = speaker_similar_exp,
              x.label = "Mindfulness", y.label = "Empathic accuracy", interval=T, int_type="confidence") 


```


