---
title: "Manuscript Analyses"
author: ""
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 2
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen=999)
```


# load packages
```{r}
library("lme4")
library("lmerTest")
library("ggplot2")
library("Rcpp")
library("lm.beta")
library("psych")
library(jtools)
library(dplyr)
library(tidyr)
library(interactions)
library(ggsci)
library(psych)
library(devtools)

```

# load and inspect data
```{r}
df <- read.csv('../data/data.csv')

df$cond <- relevel(df$cond, ref='control')


#calculate MAAS mean scores and inspect
df$maas <- rowMeans(df[,c("maas1", "maas2", "maas3", "maas4", "maas5", "maas6", "maas7", "maas8", "maas9", "maas10", "maas11", "maas12", "maas13", "maas14", "maas15")])
hist(df$maas)

#inspect data and outliers
table(scale(df$EA_corr)) 
table(scale(df$factual_accuracy)) 
table(scale(df$maas))
table(df$speaker_similar_exp)# only one person for answer 3
df<- dplyr::mutate(df, similar_combined= case_when(
  speaker_similar_exp == 1  ~ 0, 
  speaker_similar_exp == 2  ~ 1, 
  speaker_similar_exp == 3  ~ 1
))  




```


# demographics
```{r}
summary(df$age)
sd(df$age)
table (df$race)
```

# MAAS crombach's alpha
```{r}
df_maas <- df[ , c("maas1", "maas2", "maas3", "maas4", "maas5", "maas6", "maas7",
                      "maas8", "maas9","maas10", "maas11","maas12","maas13", "maas14","maas15")]    
alpha(df_maas)
```

# transform variables 
```{r}

#Fisher's r to z transformation
psych::describe(df$EA_corr)
# skew  kurtosis  se
#-0.87  1.03      0.01
df$EA_corr_raw =df$EA_corr
df$EA_corr = fisherz(df$EA_corr)

psych::describe(df$EA_corr)
#skew   kurtosis  se
#0.17   0.34      0.03

#mean-center variables
df$maas = df$maas - mean(df$maas,na.rm=T)
df$EA_corr = df$EA_corr - mean(df$EA_corr,na.rm=T)

df$similar_combined_num = df$similar_combined- mean(df$similar_combined,na.rm=T)
df$similar_combined <- as.factor(df$similar_combined_num)  
df$factual_accuracy = df$factual_accuracy - mean(df$factual_accuracy,na.rm=T)



```
# define function to make results table
```{r}
make_table = function(model) {
  require(broom)
  require(knitr)
  require(lm.beta)
  
  print(data.frame(R2 = summary(model)$r.squared,
             adjusted_R2 = summary(model)$adj.r.squared,
             df = summary(model)$df[2]) %>%
    knitr::kable(format = "pandoc", digits = 3))
    
  print(model %>%
    broom::tidy(conf.int = TRUE) %>%
    mutate(p.value = ifelse(p.value < .001, "< .001", gsub("0.(.*)", ".\\1", sprintf("%.3f", p.value))),
           `95% CI` = sprintf("%.3f, %.3f", conf.low, conf.high),
           term = gsub("as.factor", "", term)) %>%
    select(-contains("conf")) %>%
    rename("b" = estimate,
           "SE" = std.error,
           "t" = statistic,
           "p" = p.value) %>%
    select(term, b, `95% CI`, everything()) %>%
    knitr::kable(format = "pandoc", digits = 3))
}
```

# Results


# mindfulness * experience similarity --> factual accuracy
```{r}
test <-lm(factual_accuracy ~similar_combined*maas+cond,df)
make_table(test)
lm.beta(test)
```

# mindfulness * experience similarity --> empathic accuracy
```{r}
test <-lm(EA_corr ~ similar_combined*maas +cond,df)
make_table(test)
lm.beta(test)

```

# simple slopes analysis
```{r}


#+-sd for mindfulness
df$maas_low <- df$maas + sd(df$maas, na.rm=T)
df$maas_high <- df$maas - sd(df$maas, na.rm=T)

test <- lm(EA_corr~ similar_combined*maas_high+as.factor(cond), df)
make_table(test)
lm.beta(test)

test <- lm(EA_corr ~ similar_combined*maas +as.factor(cond), df)
make_table(test)
lm.beta(test)

test <- lm(EA_corr ~ similar_combined*maas_low +as.factor(cond), df)
make_table(test)
lm.beta(test)
```


# mindfulness --> empathic accuracy among subgroups of participants with vs. without similar past experience
```{r}

similar_no =subset(df, speaker_similar_exp==1)
similar_yes =subset(df, speaker_similar_exp==2 |  speaker_similar_exp == 3)

test <- lm(EA_corr~ maas+as.factor(cond), similar_no)
make_table(test)
lm.beta(test)

test <- lm(EA_corr~ maas+as.factor(cond), similar_yes)
make_table(test)
lm.beta(test)

```


# Figures
## Figure 1A
```{r}


df$Mindfulness <- df$maas
test <- lm(EA_corr~ Mindfulness * similar_combined_num +as.factor(cond), df)
interact_plot(test, pred = similar_combined_num, modx = Mindfulness,
              x.label = "Experience similarity", y.label = "Empathic accuracy", interval=T, int_type="confidence") 


```


## Figure 1B
```{r}
df3 = df %>% 
  dplyr::select("similar_combined","EA_corr", "maas") %>% 
  na.omit()

ggplot(df3) +
  aes(x = maas, y = EA_corr, color = as.factor(similar_combined)) +
  geom_point(size=1) +
  stat_smooth(method = "lm") +
  theme_bw() + theme(panel.grid.minor = element_blank())+  
  theme(panel.grid.major = element_blank()) +  
  xlab("mindfulness")  + ylab("empathic accuracy") +
  scale_color_jama()



```


```{r}
#SI2. Results using non-Fisher r to z transformed empathic accuracy scores

test <-lm(EA_corr_raw ~ similar_combined*maas+cond,df)
make_table(test)
lm.beta(test)

```

```{r}
#SI3. Treating the experience similarity as a continuous variable. 

df$similar = df$speaker_similar_exp- mean(df$speaker_similar_exp,na.rm=T)

test <-lm(factual_accuracy ~similar*maas+cond,df)
make_table(test)
lm.beta(test)

test <-lm(EA_corr ~ similar*maas+cond,df)
make_table(test)
lm.beta(test)

```


```{r}

#extra analysis
dfw <- read.csv('../data/data_windowed.csv')
dfw$rAvg = fisherz(dfw$rAvg)

dfw = dfw %>%
  group_by(pID) %>% 
  dplyr::mutate(rAvg_mean= mean(rAvg, na.rm = TRUE),
                val_mean=mean(val, na.rm = TRUE), 
                inten_mean=mean(inten, na.rm = TRUE))

#within-person mean center 
dfw$rAvg = dfw$rAvg - dfw$rAvg_mean 
dfw$val = dfw$val - dfw$val_mean 
dfw$inten = dfw$inten - dfw$inten_mean 
 #grand mean center
dfw$maas = dfw$maas - mean(dfw$maas,na.rm=T)
dfw$similar_combined = dfw$similar_combined- mean(dfw$similar_combined,na.rm=T)
dfw$similar_combined <- as.factor(dfw$similar_combined)  

#speaker's intensity and valence predicting empathic accuracy
summary(lmer(rAvg ~ val * inten + (1|pID) + (1|T), dfw))

#interaction with maas
summary(lmer(rAvg ~ val + inten + val*maas + inten*maas + (1|pID) + (1|T), dfw))

summary(lmer(rAvg ~ val + inten + val*similar_combined + inten*similar_combined+ (1|pID) + (1|T), dfw))

```
