---
title: "Purpose Craving Manuscript Analyses"
author: "Yoona Kang"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document:
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen=999)
```


# load packages
```{r}
library(mediation)
library(dplyr)
library(lme4)
library(Hmisc)
library(lmerTest) # useful if you want to get the p values for the models of each path. 
#detach("package:lmerTest", unload = TRUE) # detach and re-run models prior to attempting mediation.
library(interactions)
library(jtools)
library(ggsci)
library(weights)
library(EMAtools)
library(corrplot)
library(raster)

rm(list = ls())

```

# load and inspect data
```{r}
df_wide <- read.csv("./df_wide.csv")
df_long <- read.csv("./df_long.csv")


df_long$condition <- relevel(df_long$condition, ref='control')
length(unique(df_wide$pID)) #54=number of participants who did both fmri and ema

#define a function to get standardized beta from lmer
lmer.beta <- function(object) {
  sdy <- sd(attr(object, "resp")$y) # the y values are now in the 'y' slot 
  ###                                 of the resp attribute
  sdx <- apply(attr(object, "pp")$X, 2, sd) # And the X matriz is in the 'X' slot of the pp attr
  sc <- fixef(object)*sdx/sdy
  #mimic se.ranef from pacakge "arm"
  se.fixef <- function(obj) as.data.frame(summary(obj)[10])[,2] # last change - extracting 
  ##             the standard errors from the summary
  se <- se.fixef(object)*sdx/sdy
  return(data.frame(stdcoef=sc, stdse=se))
}


```


# METHODS

## days between fMRI and EMA
```{r}
summary(df_wide$days_between)
sd(df_wide$days_between)

```

## demographics
```{r}

#Table 1
summary(df_wide$age)
sd(df_wide$age)
table(df_wide$gender)

#women   men  Other 
#   37     16      1 
    
37/54*100 #women
16/54*100 #men
1/54*100 #other

table(df_wide$race_numeric)
summary(df_wide$race_numeric) #no NAs

# white  asian  black  latino/a  other 
# 26 16  2  3  7 

26/54*100 #white
16/54*100 #asian
2/54*100 #black
3/54*100 #latino/a
7/54*100 #other

summary(df_wide$rung_group)
sd(df_wide$rung_group, na.rm=T)

```


# RESULTS

## Descriptives

### Alcohol Descriptives
```{r}

length(df_wide$pID)
table(df_wide$hadalcohol_total) #out of 54, 11 never #had alcohol
(54-11)/54*100 # total percentage who reported having alcohol at least once
by(df_wide$hadalcohol_total, df_wide$gender, table) #0 drinks for 5F 6M 
table(df_wide$gender) #37F 16M (exclude 1 Other for percentage comparison to be consistent with the SAMHSA report)
(16-6)/16*100 #men percentage had alcohol
(37-5)/37*100 #women percentage had alcohol

#The average number of drinking occasions
summary(df_wide$hadalcohol_total) 
sd(df_wide$hadalcohol_total)
```

### Average daily alcohol use  
```{r}
### Of the ones who reported having drank alcohol at least once
drank = subset(df_long, drinks>0 )
drank = drank %>%
  group_by(pID) %>% 
  dplyr::mutate(drinks_meanWN = mean(drinks, na.rm = T))%>%
  dplyr::select(pID, drinks_meanWN) %>%
  unique()


#The average number of drinks per drinking occasion
summary(drank$drinks_meanWN)
sd(drank$drinks_meanWN)

```
### Purpose within-person variability 
```{r}

# Purpose: Calculate the within-person coefficient of variation
cv=df_long %>% 
  dplyr::group_by(pID) %>%  
  dplyr::mutate(purpose_within_cv = cv(purpose_daily, na.rm=T))%>%
  ungroup()%>%
  dplyr::select(pID, purpose_within_cv) %>%
  unique()

#remove two subjects who reported only 1 and 2 times and so have zero variability from mean calculation
cv=subset(cv, purpose_within_cv !=0)
summary(cv$purpose_within_cv)


```
### VS activity from rest [doublecheck]

```{r}
summary(df_wide$vs_alc_react)
sd(df_wide$vs_alc_react)
t.test(df_wide$vs_alc_react)

```

# Daily Purpose, Neural Alcohol Cue Reactivity, and Alcohol Craving Predicting Subsequent Alcohol Consumption

#higher ventral striatum cue reactivity was associated with greater average alcohol craving throughout the EMA period 
```{r}
cor.test(df_wide$craving_mean, df_wide$vs_alc_react)
```


## Standardize variables
```{r}
df_long$purpose_dailyz <- with(df_long, ave(purpose_daily, pID, FUN=stdz))
df_long$purpose_daily_previousz <- with(df_long, ave(purpose_daily_previous, pID, FUN=stdz))
df_long$craving_previousz <- with(df_long, ave(craving_previous, pID, FUN=stdz))
df_long$drinksz <- with(df_long, ave(drinks, pID, FUN=stdz))
```



# craving * purpose * VS --> drinking
```{r}
test <- lmer(drinksz ~ 0 + craving_previousz * purpose_daily_previousz * vs_alc_react +
              (0 + craving_previousz | site / groupID / pID) +
              (0 + purpose_daily_previousz | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition),
            df_long)
summ(test, digits = 3, confint = TRUE)
summary(test)
lmer.beta(test)
lme.dscore(test,df_long,type="lme4")

```


# Follow-up simple slopes analyses 

## Within-person mean-center variables following Aiken et al., 1991
```{r}
df_long$purpose_cen = df_long$purpose_daily_previous - df_long$purpose_mean
df_long$craving_cen = df_long$craving_previous - df_long$craving_mean
df_long$drinks_cen = df_long$drinks - df_long$drinks_mean
df_long$vs_cen = df_long$vs_alc_react - mean(df_long$vs_alc_react, na.rm=T)
df_long$age_cen = df_long$age - mean(df_long$age,na.rm=T)
df_long$rung_group_cen = df_long$rung_group - mean(df_long$rung_group, na.rm=T) 

df_long = df_long %>%
  group_by(pID) %>%
  dplyr::mutate(purpose_cen_sd = sd(purpose_cen, na.rm=T))
                
df_long$purpose_low = df_long$purpose_cen + df_long$purpose_cen_sd
df_long$purpose_high = df_long$purpose_cen - df_long$purpose_cen_sd

df_long$vs_low <- df_long$vs_cen + sd(df_long$vs_cen, na.rm=T)
df_long$vs_high <- df_long$vs_cen - sd(df_long$vs_cen, na.rm=T)
```

# simple slopes analysis
```{r}
## (SI7) see stats for vs_cen for purpose*vs two way interaction (reported stats at purpose_low and purpose_high in text) 
## see stats for vs_cen:craving_cen for purpose*vs*craving three-way interaction
test <- lmer(drinks_cen ~ 0 + purpose_low * vs_cen * craving_cen +
               (0 + purpose_low | site / groupID / pID) + 
               (0 + craving_cen | site / groupID / pID) + 
               age_cen+as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition),
             df_long, control = lmerControl(optimizer="bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

test <- lmer(drinks_cen ~ 0 + purpose_cen * vs_cen * craving_cen + 
               (0 + purpose_cen | site / groupID / pID) + 
               (0 + craving_cen | site / groupID /  pID) + 
               age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition),
             df_long, control = lmerControl(optimizer="bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

test <- lmer(drinks_cen ~ 0 + purpose_high * vs_cen * craving_cen + 
               (0 + purpose_high | site / groupID / pID) + 
               (0 + craving_cen | site / groupID / pID) +
               age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition),
             df_long, control = lmerControl(optimizer="bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

```


# Table 2
```{r}
test <- lmer(drinksz ~ 0 + craving_previousz * purpose_daily_previousz * vs_alc_react +
               (0 + craving_previousz | site / groupID / pID) +
               (0 + purpose_daily_previousz | site / groupID / pID)+
               age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition),
             df_long)
summ(test, digits = 3, confint = TRUE)
summary(test)
lmer.beta(test)
lme.dscore(test, df_long,type="lme4")

```

# Figure1
```{r}

test <- lmer(drinks_cen ~ 0 + purpose_cen * vs_cen * craving_cen + 
               (0 + purpose_cen | site / groupID / pID) +
               (0 + craving_cen | site / groupID / pID) +
               age_cen + gender + race_numeric + rung_group_cen + condition,
             df_long, control = lmerControl(optimizer="bobyqa"))

interact_plot(test, pred = craving_cen, modx=vs_cen, mod2=purpose_cen,interval = TRUE,
              x.label = "craving", y.label = "drinking",
              int.type = "confidence", int.width = 0.95)
```

# Discussion

```{r}
# the average levels of daily purpose throughout the EMA period were not significantly associated with the neural reactivity within the ventral striatum 
cor.test(df_wide$vs_alc_react, df_wide$purpose_mean)
#cor.test(df_wide$vs_alc_react, df_wide$drinks_total)
```

# Supplementary information (SI)

## SI1. Recruitment and participant sample characteristics 

```{r}
summary(df_wide$race) 
# Demographic variables were not associated with any of the main study variables (ps>.10),with an exception that men compared to women showed greater alcohol cue reactivity within the ventral striatum
v1 = df_wide %>%
  dplyr::select(gender_numeric, age, race_numeric, rung_group, purpose_mean, vs_alc_react, drinks_mean)


M <- Hmisc::rcorr(as.matrix(v1))
corrplot(M$r, p.mat = M$P, insig = "label_sig",
         sig.level = c(.001, .05, .10), pch.cex=0.5, pch.col = "white",
         method="color", type="lower")

summary(aov(purpose_mean ~ gender, df_wide))
summary(aov(vs_alc_react ~ gender, df_wide)) #sig
summary(aov(drinks_mean ~ gender, df_wide))

summary(aov(purpose_mean ~ race_numeric, df_wide))
summary(aov(vs_alc_react ~ race_numeric, df_wide))
summary(aov(drinks_mean ~ race_numeric, df_wide))

men=subset(df_wide, gender=="Male")
women=subset(df_wide, gender=="Female")

summary(men$vs_alc_react)
summary(women$vs_alc_react)

```
##SI3. Ecological momentary assessment (EMA) data collection and transformation [doublecheck]

```{r}

#missing data
length(unique(df_wide$vs_alc_react)) # 54 out of 54; none missing
summary(df_long$craving_previousz) #616
summary(df_long$purpose_daily_previousz) #495
summary(df_long$drinksz) #1042

```

## SI4. Neurosynth map of ‘craving’
```{r}

t.test(df_wide$neurosynth_craving_alc_react)

test <- lmer(drinksz ~ 0 + craving_previousz * purpose_daily_previousz * neurosynth_craving_alc_react +
               (0 + craving_previousz | site / groupID / pID) +
               (0 + purpose_daily_previousz | site / groupID / pID) + 
               age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition),
             df_long)
summ(test, digits = 3, confint = TRUE)
summary(test)
lmer.beta(test)
lme.dscore(test, df_long,type = "lme4")


#followup simple slopes analyses
df_long$neurosynth_craving_alc_react_cen = df_long$neurosynth_craving_alc_react - mean(df_long$neurosynth_craving_alc_react, na.rm=T)

test <- lmer(drinks_cen ~ 0 + purpose_low * neurosynth_craving_alc_react_cen * craving_cen +
               (0 + purpose_low | site / groupID / pID) +
               (0 + craving_cen | site / groupID / pID) +
               age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition),
             df_long, control = lmerControl(optimizer="bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

test <-lmer(drinks_cen ~ 0 + purpose_cen * neurosynth_craving_alc_react_cen * craving_cen  +
              (0 + purpose_low | site / groupID / pID) +
              (0 + craving_cen | groupID / pID) + 
              age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition),
            df_long, control = lmerControl(optimizer="bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

test <- lmer(drinks_cen ~ 0 + purpose_high * neurosynth_craving_alc_react_cen * craving_cen +
               (0 + purpose_low | site / groupID / pID) +
               (0 + craving_cen | site / groupID / pID) +
               age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition), 
             df_long, control = lmerControl(optimizer = "bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

```
## Figure SI.
```{r}
test <- lmer(drinks_cen ~ 0 + purpose_cen * neurosynth_craving_alc_react_cen * craving_cen +
               (0 + purpose_cen | site / groupID / pID) +
               (0 + craving_cen | site / groupID / pID) + 
               age_cen + gender + race_numeric + rung_group_cen + condition,
             df_long, control = lmerControl(optimizer = "bobyqa"))

interact_plot(test, pred = craving_cen, modx = neurosynth_craving_alc_react_cen, mod2 = purpose_cen, 
              interval = TRUE,
              x.label = "craving", y.label = "drinking",
              int.type = "confidence", int.width = 0.95)


```

## SI5. Multilevel regression model fit and test of multicollinearity

### AIC

```{r}

AIC(lmer(drinksz ~ 0 + craving_previousz * purpose_daily_previousz * vs_alc_react +
              (0 + craving_previousz | site / groupID / pID) +
              (0 + purpose_daily_previousz | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition),
            df_long))

AIC(lmer(drinksz ~ 0 + craving_previousz *  vs_alc_react +
              (0 + craving_previousz | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition),
            df_long))

AIC(lmer(drinksz ~ 0 + purpose_daily_previousz * vs_alc_react +
              (0 + purpose_daily_previousz | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition),
            df_long))

AIC(lmer(drinksz ~ 0 + craving_previousz * purpose_daily_previousz +
              (0 + craving_previousz | site / groupID / pID) +
              (0 + purpose_daily_previousz | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition),
            df_long))


```


### VIF 

```{r}
vif.lme <- function (fit) {
  ## adapted from rms::vif
  v <- vcov(fit)
  nam <- names(fixef(fit))
  ## exclude intercepts
  ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
  if (ns > 0) {
    v <- v[-(1:ns), -(1:ns), drop = FALSE]
    nam <- nam[-(1:ns)] }
  d <- diag(v)^0.5
  v <- diag(solve(v/(d %o% d)))
  names(v) <- nam
  v }

test <- lmer(drinksz ~ 0 + purpose_daily_previousz * vs_alc_react * craving_previousz +
               (0 + purpose_daily_previousz | site / groupID / pID) + 
               (0 + craving_previousz | site / groupID / pID),
             df_long)

vif <- data.frame(vif.lme(test)) #low correlations of vif=~1
summary(vif$vif.lme.test.)
```

### FDR correction [doublecheck]
```{r}
summary(lmer(drinks_cen ~ 0 + purpose_low * vs_cen * craving_cen +
               (0 + purpose_low | site / groupID / pID) + 
               (0 + craving_cen | site / groupID / pID) + 
               age_cen+as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition),
             df_long, control = lmerControl(optimizer="bobyqa")))

summary(lmer(drinks_cen ~ 0 + purpose_cen * vs_cen * craving_cen + 
               (0 + purpose_cen | site / groupID / pID) + 
               (0 + craving_cen | site / groupID /  pID) + 
               age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition),
             df_long, control = lmerControl(optimizer="bobyqa")))

summary(lmer(drinks_cen ~ 0 + purpose_high * vs_cen * craving_cen + 
               (0 + purpose_high | site / groupID / pID) + 
               (0 + craving_cen | site / groupID / pID) +
               age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition),
             df_long, control = lmerControl(optimizer="bobyqa")))

p = c(0.00559,0.03210,0.22167)
round(p.adjust(p, "BH"), 3)

```


## SI6. Results not controlling for potential covariates

```{r}
test <- lmer(drinksz ~ 0 + craving_previousz * purpose_daily_previousz * vs_alc_react +
               (0 + craving_previousz | site / groupID / pID) + 
               (0 + purpose_daily_previousz | site / groupID / pID),
             df_long)
summ(test, digits = 3, confint = TRUE)
summary(test)
lmer.beta(test)


test <- lmer(drinks_cen ~ 0 + purpose_low * vs_cen * craving_cen + 
               (0 + purpose_low | site / groupID / pID) + 
               (0 + craving_cen | site / groupID / pID),
             df_long)
summ(test, digits = 3, confint = TRUE)
summary(test)

test <- lmer(drinks_cen ~ 0 + purpose_cen * vs_cen * craving_cen + 
               (0 + purpose_cen | site / groupID / pID) + 
               (0 + craving_cen | site / groupID / pID),
             df_long,control = lmerControl(optimizer = "bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

test <- lmer(drinks_cen ~ 0 + purpose_high * vs_cen * craving_cen + 
               (0 + purpose_high | site / groupID / pID) +
               (0 + craving_cen | site / groupID / pID), df_long)
summ(test, digits = 3, confint = TRUE)
summary(test)

```



## SI8. Results controlling for the number of days between the fMRI and EMA data collection.

```{r}

#correlations between days and main variables. 

v1 = df_wide %>%
  dplyr::select(days_between, purpose_mean, drinks_total, vs_alc_react)

M <- Hmisc::rcorr(as.matrix(v1))

corrplot(M$r, p.mat = M$P, insig = "label_sig",
         sig.level = c(.001, .05, .10), pch.cex=0.5, pch.col = "white",
         method="color", type="lower")

test <- lmer(drinksz ~ 0 + craving_previousz * purpose_daily_previousz * vs_alc_react +
              (0 + craving_previousz | site / groupID / pID) +
              (0 + purpose_daily_previousz | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition) + days_between,
            df_long)
summ(test, digits = 3, confint = TRUE)
summary(test)


test <- lmer(drinks_cen ~ 0 + purpose_low * vs_cen * craving_cen +
               (0 + purpose_low | site / groupID / pID) + 
               (0 + craving_cen | site / groupID / pID) + 
               age_cen+as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition) + days_between,
             df_long, control = lmerControl(optimizer="bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

test <- lmer(drinks_cen ~ 0 + purpose_cen * vs_cen * craving_cen + 
               (0 + purpose_cen | site / groupID / pID) + 
               (0 + craving_cen | site / groupID /  pID) + 
               age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition) + days_between,
             df_long, control = lmerControl(optimizer="bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

test <- lmer(drinks_cen ~ 0 + purpose_high * vs_cen * craving_cen + 
               (0 + purpose_high | site / groupID / pID) + 
               (0 + craving_cen | site / groupID / pID) +
               age_cen + as.factor(gender) + as.factor(race_numeric) + rung_group_cen + as.factor(condition) + days_between,
             df_long, control = lmerControl(optimizer="bobyqa"))
summ(test, digits = 3, confint = TRUE)
summary(test)

```

# R1.5. scanner craving scores
```{r}

inscan <- read.csv( "~/Box Sync/MURI_Data_830189/data/EMA_Round1/Yoona_PurposeAnalysis/data/source/inscan_crave_deided.csv") 
summary(inscan$alc_react_crave)
sd(inscan$alc_react_crave)

df2 <-merge(df_wide,inscan,by="pID", all=F)
summary(lm(craving_mean ~ alc_react_crave + condition,df2))


df3=df2 %>% 
  dplyr::select(pID, alc_react_crave)

df_long <-merge(df_long,df3,by="pID", all=T)


test <- lmer(drinksz ~ 0 + alc_react_crave * purpose_daily_previousz * vs_alc_react +
              (0 + purpose_daily_previousz | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition),
            df_long)
summary(test)

```



#SME4.quadratic trends

```{r}
#create quadratic and cubic terms

df_long$cravingSqu = df_long$craving_previousz * df_long$craving_previousz
df_long$cravingCu = df_long$craving_previousz * df_long$craving_previousz * df_long$craving_previousz

test <- lmer(drinksz ~ 0 + craving_previousz + cravingSqu +
              (0 + craving_previousz | site / groupID / pID) +
              (0 + cravingSqu | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition) + days_between,
            df_long)
summ(test, digits = 3, confint = TRUE)
AIC(test)



test <- lmer(drinksz ~ 0 + craving_previousz + cravingSqu +
              (0 + craving_previousz | site / groupID / pID) +
              (0 + cravingSqu | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition) + days_between,
            df_long)
summ(test, digits = 3, confint = TRUE)
AIC(test)


test <- lmer(drinksz ~ 0 + craving_previousz + cravingCu +
              (0 + craving_previousz | site / groupID / pID) +
              (0 + cravingCu | site / groupID / pID) +
              age + as.factor(gender) + as.factor(race_numeric) + rung_group + as.factor(condition) + days_between,
            df_long)
summ(test, digits = 3, confint = TRUE)
AIC(test)

```

##SIx
```{r}


```
